---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pearson.diagram

<!-- badges: start -->
<!-- badges: end -->

The Pearson Diagram R package provides tools to visualize distributions using Pearson Diagrams. It helps users calculate skewness and kurtosis for unknown data distributions, plot the Pearson Diagram, and compare multiple distributions. This package leverages **Rcpp** for efficient computation and follows the **S3** object-oriented programming paradigm in R.

## Key Features

- **Plotting Pearson Diagrams:** Visualize distributions on a Pearson Diagram by plotting skewness and kurtosis.
- **Comparison of Multiple Distributions:** Compute and compare skewness and kurtosis for multiple distributions.
- **Efficient Calculation with Rcpp:** Uses C++ for high-performance calculation of skewness and kurtosis.
- **S3 Object-Oriented Programming:** Encapsulates functionality within S3 classes for Pearson Diagrams.

## Installation

You can install the development version of pearson.diagram like so:


``` r
# Yet to be added
```

## Example

Work on the below command to check if the package is build successfully.
```{r}
devtools::check()
devtools::check(document = TRUE, manual = TRUE)
```
