---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pearson.diagram

<!-- badges: start -->
<!-- badges: end -->

The `pearson.diagram` package distinguishes different distribution by visualizing their skewness and kurtosis on a Pearson diagram. It helps users calculate skewness and kurtosis for unknown data distributions, plot the Pearson Diagram, and compare multiple distributions. This package leverages **Rcpp** for efficient computation and follows the **S3** object-oriented programming paradigm in R. This file provides a detailed guide on how to use the package, explains its backend functionality, and includes a test plan that can be followed to verify the package’s accuracy.

## Key Features

- **Display Pearson Diagrams:** Use a Pearson Diagram to view several distributions by showing their skewness and kurtosis values for easy comparison.

- **Plotting Unknown Data:** Allows users to calculate and plot skewness and kurtosis, which helps in the identification of the true distribution.

- **Bootstrapping Functionality:** (A placeholder for future development) Users can generate bootstrap samples from their data and plot variations of skewness and kurtosis on the diagram.

- **Efficient Calculation with Rcpp:** Rcpp uses C++ to carry out high-performance skewness and kurtosis calculations, enabling efficient processing of huge datasets.

- **Hovering Functionality:** The Interactive Hovering Feature allows for user interactions by offering customized hover tooltips and advanced visual feedback when interacting with Pearson Diagram points.

- **S3 Object-Oriented Programming:** Implements functionality using S3 classes, encapsulating Pearson Diagrams as objects and providing methods to add data points and plot the results in a structured, object-oriented way.

## 1. Installation and loading the package
Install the package from GitHub using the following command:

``` {r}
# devtools::install_github("renswickd/R-Package-for-Pearson-Diagram")
```
Once installed, load the package:
```{r}
library(pearson.diagram)
```

## 2. Overview of Main Functions
### 2.1 `plot_diagram()` (Main Function)
The Pearson diagram illustrates the relationship between the skewness and kurtosis of several probability distributions. The $x-\text{axis}$ shows the square of skewness, while the $y-\text{axis}$ shows the kurtosis. This figure helps in visually identifying the properties of various distributions and how they fit in the skewness-kurtosis space.

  Skewness $\beta_{1}=\mu_{3}^{2}/\mu_{2}^{3}$
  
  Kurtosis $\beta_{2}=\mu_{4}/\mu_{2}^{2}$

* This is the main function that generates the Pearson diagram. It accepts a variety of input parameters, including a Pearson object, custom input data, and option to enable bootstrap functionality or hover-over interactivity.
* If custom (unknown) data is provided, it calculates the square of skewness and kurtosis for the data points and highlights them on the plot. Otherwise, the function generates a pearson plot with known distribution families and no extra points.

```{r, out.width="50%"}
pd <- PearsonDiagram() # Create empty pearson diagram object
data <- rnorm(10000) # Generate a random normal dataset
plot_diagram(pd, input_data = data) # Plot the Pearson diagram with this single dataset
```


#### 2.1.1 `canvas_creation()`
The `canvas_creation()` function builds the layout of the Pearson diagram. It contains known distribution families with three different sorts of representations:

  * *Point Representation:* Normal and Uniform distributions are represented by a single point since their skewness and kurtosis values are fixed.

  * *Line Representation:* Exponential and Gamma distributions, whose skewness and kurtosis vary with a parameter, are displayed as lines.

  * *Area Representation:* In Beta distributions, where multiple shape parameters influence skewness and kurtosis, an area can be seen to show a range of skewness and kurtosis values.
  
```{r, out.width="50%"}
canvas_creation()
```

### 2.2 `cpp_calculate_moments()`
This function calculates the skewness and kurtosis using high-performance Rcpp code. It is optimized to handle large datasets efficiently. It returns a list containing the square of skewness and kurtosis.

```{r}
moments <- cpp_calculate_moments(data)
print(moments)
```

### 2.3 `Object Oriented Programming`
The `pearson.diagram` package utilizes OOP approach using R S3 classes. This gives for greater flexibility when it comes to adding custom data points and customizing plot characteristics. By designing the package in such a way, it is easier to expand or modify the behavior of each function for different kinds of data and objects. The following functions are crucial for implementing OOP:

  * `PearsonDiagram()`: Creates an object of the `PearsonDiagram` class containing information on values required to create the canvas and plot the points such as square of skewness, kurtosis, and the distribution name.

  * `add_point.PearsonDiagram()`: Uses the dispatch function to add points to the `PearsonDiagram` object. This method allows users to add additional data to the object in a systematic manner.

  * `plot_diagram.PearsonDiagram()` plots the Pearson diagram by dispatching based on the object's class.


### 2.4 `add_point()`
The function`add_point()` inserts a point to the PearsonDiagram object. This point is distinguished by the square of skewness, kurtosis, and the optional distribution name. The point may indicate known distributions or custom data provided by the user.

#### `Backend Functioning`:

  * For given distributions (5 distributions mentioned in section[2.1]), analytical formulae are used to calculate the square of skewness and kurtosis.
  * For unknown input data, the function uses the `cpp_calculate_moments()` function to efficiently compute these values.

### 2.5 `validate_input()`
The `validate_input()` function ensures that the user-provided data is valid. The function eliminates runtime issues by requiring valid input and ensuring accurate skewness and kurtosis computations. Below are the conditions validated by the function:

   - *Numeric:* The input data must be numeric.
   - *Length:* The input must contain at least 3 data points to allow for meaningful calculations of skewness and kurtosis.
   - *Missing values:* The input must not contain NA or NaN values.
   - *Infinite values:* The input must not contain Inf or -Inf values.
   - *Zero variance:* The input data must have non-zero variance. Constant data will not provide meaningful statistics for skewness or kurtosis.
   
### 2.6 Customization (`highlight_point_on_hover()`)
This function is currently implemented as a PLACEHOLDER, but upcoming versions will have extensive interactivity with the Pearson diagram.

#### Planned Features:

  * Users will be able to hover over points in the diagram to see tooltips with information such as skewness, kurtosis, and the distribution name. Also, when user hover over various portions of the plot, user can trigger actions like zooming in.
  * Beyond hovering over the plotted points, I aim to implement the plot customization (including title, font, color etc.)
  * Aiming to enable the users to download the plot to specified directory in local.
  
### 2.7 Bootstrapping Functionality
The `plot_diagram()` method has a bootstrapping option, which, if enabled, bootstraps the input data. Bootstrapping is the process of resampling input data using replacement to produce a large number of simulated samples. The resampled datasets are then used to estimate the skewness and kurtosis distributions.

**Visualization:** Bootstrapped samples will be provided alongside the main data points, allowing users to view the variability and distribution of skewness and kurtosis for their input data.

## 3. Example Workflow
Here’s a complete example workflow to help you get started with the `pearson.diagram` package: Detailed examples and steps explained are available in the vignettes of the package.

```{r, out.width="50%"}
pd <- PearsonDiagram()
data1 <- rnorm(1000)            # Normal distribution
data2 <- rgamma(1000, shape = 2) # Gamma distribution

# Plot Pearson diagram with multiple datasets
plot_diagram(pd, input_data = list(data1, data2))
```

## 4. Test Plan
Testing the `pearson.diagram` package is critical to ensuring that its functionality works as expected, including basic functions such as calculating skewness and kurtosis, creating visualizations, processing user inputs, and maintaining the package's performance under various conditions. The test plan is divided into two sections: implemented tests and yet-to-be implemented tests.

### 4.1 Unit Testing
##### 4.1.1 `plot_diagram()` - Validate that the function correctly plots the Pearson diagram with the correct visual representation of known and custom distributions.
  * *Test Input:* A `PearsonDiagram` object with points representing Normal, Exponential, and Gamma distributions.
  * *Expected Result:* Correct plot with distinct representations (point, line, area) for each distribution.
  * **Status: Implemented**
```{r}
# refer (section [2.1], section [3])
```

##### 4.1.2 cpp_calculate_skewness_kurtosis() - Ensure that this function accurately calculates skewness and kurtosis for known distributions.
  * *Test Input:* A vector of random data from a standard normal distribution.
  * *Expected Result:* Skewness close to 0 and kurtosis close to 3.
  * **Status: Implemented**
```{r}
# refer (section [2.2])
```

##### 4.1.3 add_point() - Ensure that a point representing a distribution (given its skewness and kurtosis) is correctly added to the PearsonDiagram object.
  * *Test Input:* Skewness and kurtosis values for a known distribution like Gamma.
  * *Expected Result:* Point added to the PearsonDiagram object with accurate properties.
  * **Status: Implemented**
  
##### 4.1.4 validate_input() - Ensure that the function correctly validates input datasets and raises appropriate error messages for invalid inputs (e.g., non-numeric data, missing values, infinite values).
  * *Test Input:* A vector / list containing NA values.
  * *Expected Result:* Error message indicating invalid input.
  * **Status: Implemented**

### 4.2 Edge Case Testing
##### 4.2.1 validate_input() - Validate that the function handles edge cases such as datasets with outliers or insufficient points.
  * *Test Input:* A constant numeric vector/list.
  * *Expected Result:* Error message indicating that the data has zero variance.
  * **Status: Implemented**
  
##### 4.2.2 `plot_diagram()` - Test edge cases where the PearsonDiagram object has no points or insufficient data.
  * *Test Input:* An empty PearsonDiagram object.
  * *Expected Result:* Error message stating that there are no points to plot.
  * **Status: Implemented**
  
### 4.3 Performance Testing
##### 4.3.1 cpp_calculate_skewness_kurtosis() - Benchmark the performance of this function when processing large datasets, especially with C++ backend through Rcpp.
  * *Test Input:* Large numeric datasets with 10,000+ observations.
  * *Expected Result:* Fast and efficient calculation of skewness and kurtosis.
  * **Status: Yet to be Implemented**

##### 4.3.2 Stress Testing - Evaluate the behavior of the package under stress by processing very large or complex datasets.
  * *Test Input:* A combination of large datasets with varied distributions (Normal, Exponential, Gamma).
  * *Expected Result:* Identify performance bottlenecks, memory usage issues, or potential optimizations.
  * **Status: Yet to be Implemented**
  
##### 4.3.3 `plot_diagram()` with Bootstrapping - Test the bootstrapping functionality to ensure that it correctly visualizes resampled distributions.

  * *Test Input:* Small numeric dataset with bootstrap = TRUE.
  * *Expected Result:* Bootstrapped samples visualized on the Pearson diagram with correct representation of distribution variability.
  * **Status: Yet to be Implemented**

### 4.4 Continuous Integration (Automated Testing)
##### 4.4.1 Objective: Implement continuous integration (CI) using tools like testthat and set up pipelines for automatic testing of the package upon each commit or pull request.
  * Expected Outcome: CI will run unit tests automatically and flag any issues caused by new code changes.
  * **Status: Yet to be implemented**
