---
title: "Pearson Diagram"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pearson Diagram}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Vignette for Pearson Diagram Package
# Pearson Diagram: Skewness and Kurtosis Visualization of Distributions
The `pearson.diagram` package provides a simple and effective way to visualize the skewness and kurtosis of known and custom distributions on the Pearson diagram. This vignette demonstrates how to use the package, generate diagrams, add custom data points, and utilize the bootstrap functionality.

## Installation
If the package is not yet installed, you can install it using:

```
install_github("renswickd/R-Package-for-Pearson-Diagram")
```
## Loading the Package
```{r setup}
library(pearson.diagram)
```

## 1. Adding Input Data Points
You can add custom input data to the diagram using the `plot_diagram()` function. The input can be a single numeric vector or a list of numeric vectors, which will be plotted as points on the Pearson diagram.

### 1.1 Single Dataset Example
Let’s first add a normally distributed dataset and visualize its skewness and kurtosis:
```{r}
data <- rnorm(1000) # Generate a random normal dataset
plot_diagram(input_data = data) # Plot the Pearson diagram with this single dataset
```

### 1.2 Multiple Datasets Example
You can also provide multiple datasets (e.g., Normal, Gamma), and the function will calculate and plot each one:
```{r}
data1 <- rnorm(1000)            # Normal distribution
data2 <- rgamma(10000, shape = 2) # Gamma distribution1
data3 <- rgamma(1000, shape = 1)   # Gamma distribution2

# Plot the Pearson diagram with multiple datasets
plot_diagram(input_data = list(data1, data2, data3))
```
## 2. Using cpp_calculate_moments to Calculate Skewness and Kurtosis
For custom datasets, you can use `cpp_calculate_moments()` to efficiently compute the skewness and kurtosis using a high-performance Rcpp implementation. Here's an example of using this function on a randomly generated dataset:
```{r}
moments <- cpp_calculate_moments(data1)
print(moments)
```
## 3. Bootstrap Functionality
The `plot_diagram(`) function also supports a bootstrap argument, which allows for bootstrapping the samples and calculate skewness and kurtosis for each samples when enabled. Although the current implementation has a placeholder for this functionality, it can be extended in future versions.
```{r}
data <- rnorm(1000) # Generate a random normal dataset
plot_diagram(input_data = data, bootstrap = TRUE)
```


## 4. Outlier detection 
A basic level of identifying and providing appropriate warning regarding the extreme values of the input data will be developed in the further versions of the package.
```{r}
data1 <- c(rnorm(1000), rnorm(2, 10, 3))
plot_diagram(data1, hover=TRUE, bootstrap = TRUE, treat.outliers = TRUE)
```


